The increasing cost of happiness

### 2020-09-08 src/The increasing cost of happiness.Rmd
Add figure

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE,
                      fig.path = "../results/")

@@

## Abstract
A fundamental question for governments and people is how much happiness does a dollar buy? The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing wealth. While the location of this _change point_ has been determined, and the cost of happiness reportedly ranges between USD\$60-95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time. We use household economic data from Australia between 2002-2016 to test the relationship with both happiness and life satisfaction. We discovered no change point existed for satisfaction, but the cost of happiness has increased over those 16 years faster than inflation (i.e., cost of living).

@@

_Model selection_  
We compared the fit of the "broken-stick" model with a simpler linear model ("unbroken"), adjusted for degrees of freedom by the Bayesian Information Criterion (BIC). BIC was chosen over AIC or LRT because it will penalize our more complex model (the broken stick) more heavily. Following Raftery, differences in BIC greater than 10 were taken as "very strong evidence"; greater than 5 were "strong evidence"; greater than 2 were "positive evidence"; and less than 2 (but greater than zero) were "weak evidence" for the model with the smallest BIC ( [Raftery, 1995](https://www.sciencedirect.com/topics/pharmacology-toxicology-and-pharmaceutical-science/bayesian-information-criterion){target="_blank"}).

@@

Covariates included in subsequent sensitivity analyses were: age, age^2^, years of education, and dummy variables for partnered, single (never married), and unemployed.

@@

@@

- Cognitive wellbeing (satisfaction) had a relatively linear, positive relationship wtih household disposable income. The inflection point was equally positive and negative across years, without a uniform pattern. 
- By contrast, the relationship between affective wellbeing (happiness) and disposable income tended to be a positive but decreasing ("*broken*") function with a uniform inflection point.

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(bayesplot)

knitr::opts_chunk$set(eval = TRUE,
                      echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../figures/")

@@

## Abstract
A fundamental question for governments and people is how much happiness does a dollar buy? The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing wealth. While the location of this _change point_ has been determined, and the cost of happiness reportedly ranges between USD\$60-95K, there has been no investigation as to whether the cost of happiness has increased or decreased over time. We tested the relationship between money and both happiness and life satisfaction using household economic data from Australia between 2002-2016. We discovered no change point with wealth existed for satisfaction, but the cost of happiness has increased over those 16 years faster than inflation (i.e., cost of living). Results such as these suggest we need to understand why the cost of happiness has increased, and to consider whether policy-goals to improve wellbeing by increasing wealth are feasible.

@@

_Model selection_  
The [Widely Applicable Information Criterion (WAIC) was used for model selection](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/overfitting-regularization-and-information-criteria.html#the-problem-with-parameters){target="_blank"}. WAIC was defined as:

WAIC = -2(lppd - *p*~WAIC~)

Where lppd (log pointwise predictive density) is the total across observations of the log of the average likelihood of each observation, and *p*~WAIC~ is the effective number of free parameters determined by the sum of the variance in log-likelihood for each observation (*i*).

@@

Age (and age^2^) and education were included as covariates, and students were removed.

@@

##### Bayesian change point results

Shaded areas to the right of the vertical dotted line are credibly (95%) larger than the base year (2002).

```{r happy_change_time, fig.width = 4.9, fig.height = 9}
hifdip_gh9_fixed_fit <- read_rds("../results/hifdip_gh9_fixed_full.RDS")

as.data.frame(hifdip_gh9_fixed_fit) %>%
  select(b_alpha_Intercept, b_alpha_year2006, b_alpha_year2010, 
         b_alpha_year2014, b_alpha_year2018) %>%
  transmute(
    fy2002 = b_alpha_Intercept,
    fy2006 = mean(b_alpha_Intercept) + b_alpha_year2006,
    fy2010 = mean(b_alpha_Intercept) + b_alpha_year2010,
    fy2014 = mean(b_alpha_Intercept) + b_alpha_year2014,
    fy2018 = mean(b_alpha_Intercept) + b_alpha_year2018
  ) %>%
  mutate_all(.funs = ~brms::inv_logit_scaled(., 0, 150)*10) -> d1

color_scheme_set("red")
mcmc_areas(
  d1,
  pars = rev(colnames(d1)),
  prob = 0.95, # 80% intervals
  prob_outer = 1, # 99%
  point_est = "mean"
  ) +
  theme_test() +
  geom_vline(aes(xintercept = mean(d1$fy2002)), color = "gray", linetype = 3) +
  labs(
    subtitle = "Posterior change points (mean & 95% distribution)",
    x = "Household wealth ($000s)")
```

@@

- The change point for happiness (affective wellbeing) shifts to the right over time and requires greater household wealth.
